`r glue::glue('Initial inferCNV results for {params$library_id}')`

### 2025-04-15 analyses/infercnv-consensus-cell-type/template-notebooks/ewings-infercnv-results.Rmd
clean up some redundant data frames

celltype_df <- readr::read_tsv(ewings_tsv) |>
  dplyr::select(barcodes, consensus_annotation, ewing_annotation)

@@

# Read SCE and save UMAP coordinates
umap_df <- readr::read_rds(sce_file) |>
  scuttle::makePerCellDF(use.dimred = "UMAP") |>
  dplyr::select(
    barcodes,
    UMAP1 = UMAP.1,
    UMAP2 = UMAP.2
  )
```

@@

```{r}
# First, construct data frame of total `has_cnv` counts, with other variables needed for plots
has_cnv_df <- infercnv_df |>
  dplyr::select(
    barcodes,
    normal_reference,
    starts_with("has_cnv_")
  ) |>
  tidyr::pivot_longer(
    starts_with("has_cnv_"),
    names_to = "chr",
    values_to = "cnv"
  ) |>
  # sum all values for each cell, but separately for each run of inferCNV (aka each normal reference)
  dplyr::group_by(barcodes, normal_reference) |>
  dplyr::summarize(total_cnv_per_cell = sum(cnv)) |>
  dplyr::ungroup() |>
  # add indicator columns for whether the cell is reference or query
  dplyr::mutate(
    immune_reference_cell = barcodes %in% immune_ref_barcodes,
    subset_reference_cell = barcodes %in% subset_ref_barcodes
  ) |>
  # join in other data
  dplyr::left_join(celltype_df) |>
  dplyr::left_join(umap_df)
```

@@

```{r fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
# drop NAs from reference cells from other libraries
query_has_cnv_df <- has_cnv_df |>
  tidyr::drop_na(ewing_annotation, consensus_annotation)

@@

```{r, fig.width = 6}
query_has_cnv_df |>
  # arbitrary filtering, this is just data
  dplyr::filter(normal_reference == "all-immune") |>
  ggplot() +
  aes(x = UMAP1, y = UMAP2, color = forcats::fct_lump(ewing_annotation, 8)) +
  geom_point(size = 0.1, alpha = 0.3) +
  facet_wrap(vars(normal_reference)) +
  labs(color = "annotation") +
  guides(color = guide_legend(override.aes = list(size = 2, alpha = 1))) +
  umap_theme
```

@@

```{r, fig.height=6}
compare_refs <- has_cnv_df |>
  # bring in library_id so we can only include those cells.
  dplyr::inner_join(
    infercnv_df |>
      dplyr::select(barcodes, library_id) |>
      unique()
  ) |>
  dplyr::filter(library_id == params$library_id) |>
  dplyr::select(
    barcodes, normal_reference, total_cnv_per_cell
  ) |>
  tidyr::pivot_wider(
    names_from = normal_reference,
    values_from = total_cnv_per_cell
  ) |>
  # color points this way rather than binning to ensure small bins aren't left out
  dplyr::add_count(`all-immune`, `subset-immune`)

@@

```{r, fig.width = 10}
# bring in cluster and library information
cluster_cnv_df <- has_cnv_df |>
  dplyr::inner_join(
    infercnv_df |>
      dplyr::select(barcodes, library_id, subcluster, normal_reference) |>
      unique(),
    by = c("barcodes", "normal_reference")
  )

ggplot(cluster_cnv_df) +
  # keep in default order so "reference" and "unknown" are grouped
  aes(x = subcluster) +
  geom_bar() +
  facet_wrap(
    vars(normal_reference),
    scales = "free_x"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

```{r fig.width=12}
cluster_cnv_df |>
  # remove cells from other libraries
  dplyr::filter(library_id == params$library_id) |>
  ggplot() +
  aes(x = subcluster, fill = forcats::fct_lump(consensus_annotation, 7)) +
  geom_bar() +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap(
    vars(normal_reference),
    scales = "free_x"
  ) +
  labs(
    title = "Consensus cell types",
    fill = "annotation"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r fig.width=12}
cluster_cnv_df |>
  # remove cells from other libraries
  dplyr::filter(library_id == params$library_id) |>
  ggplot() +
  aes(x = subcluster, fill = forcats::fct_lump(ewing_annotation, 7)) +
  geom_bar() +
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(
    vars(normal_reference),
    scales = "free_x"
  ) +
  labs(
    title = "Ewing cell types",
    fill = "annotation"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

```{r fig.width=10}
all_reference_clusters <- cluster_cnv_df |>
  dplyr::filter(
    normal_reference == "all-immune",
    immune_reference_cell
  )
all_plot <- ggplot(all_reference_clusters) +
  aes(x = subcluster, fill = library_id) +
  geom_bar() +
  labs("All immune reference") +
  theme(legend.position = "bottom")

subset_reference_clusters <- cluster_cnv_df |>
  # bring in library_id
  dplyr::inner_join(
    infercnv_df |>
      dplyr::select(barcodes, library_id) |>
      unique()
  ) |>
  dplyr::filter(
    normal_reference == "subset-immune",
    subset_reference_cell
  )
subset_plot <- ggplot(subset_reference_clusters) +
  aes(x = subcluster, fill = library_id) +
  geom_bar() +
  labs("Subset immune reference") +
  theme(legend.position = "bottom")

@@

```{r}
# How many CNV per cell per cluster?
ggplot(cluster_cnv_df) +
  aes(x = subcluster, y = total_cnv_per_cell) +
  geom_point(size = 2) +
  facet_wrap(
    vars(normal_reference),
    scales = "free_x"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

>>>>

@@

# Read SCE and save UMAP coordinates and cell types
umap_df <- readr::read_rds(sce_file) |>
  scuttle::makePerCellDF(use.dimred = "UMAP") |>
  dplyr::select(
    barcodes,
    UMAP1 = UMAP.1,
    UMAP2 = UMAP.2
  ) |>
  # TODO: We currently are bringing in cell types from separate files, but in the next
  # data release they will be in the SCE objects
  dplyr::inner_join(
    readr::read_tsv(ewings_tsv) |>
      dplyr::select(barcodes, consensus_annotation, ewing_annotation)
  )
```

@@

```{r}
# First, construct data frame of total `has_cnv` counts, with other variables needed for plots
has_cnv_df <- infercnv_df |>
  dplyr::select(
    barcodes,
    normal_reference,
    starts_with("has_cnv_")
  ) |>
  tidyr::pivot_longer(
    starts_with("has_cnv_"),
    names_to = "chr",
    values_to = "cnv"
  ) |>
  # sum all values for each cell, but separately for each run of inferCNV (aka each normal reference)
  dplyr::group_by(barcodes, normal_reference) |>
  dplyr::summarize(total_cnv_per_cell = sum(cnv)) |>
  dplyr::ungroup() |>
  # add indicator columns for whether the cell is reference or query
  dplyr::mutate(
    immune_reference_cell = barcodes %in% immune_ref_barcodes,
    subset_reference_cell = barcodes %in% subset_ref_barcodes
  ) |>
  # join in SCE information
  dplyr::left_join(umap_df) |>
  # finally, bring back in cluster and library information
  dplyr::left_join(
    infercnv_df |>
      dplyr::select(barcodes, library_id, subcluster, normal_reference) |>
      unique(),
    by = c("barcodes", "normal_reference")
  )
```

@@

```{r fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
# drop cells from other libraries
query_has_cnv_df <- has_cnv_df |>
  dplyr::filter(library_id == params$library_id)

@@

```{r, fig.width = 6}
ggplot(umap_df) +
  aes(x = UMAP1, y = UMAP2, color = forcats::fct_lump(ewing_annotation, 8)) +
  geom_point(size = 0.1, alpha = 0.3) +
  labs(color = "ewings annotation") +
  guides(color = guide_legend(override.aes = list(size = 2, alpha = 1))) +
  umap_theme
```

@@

```{r, fig.height=6}
compare_refs <- has_cnv_df |>
  dplyr::filter(library_id == params$library_id) |>
  dplyr::select(
    barcodes, normal_reference, total_cnv_per_cell
  ) |>
  tidyr::pivot_wider(
    names_from = normal_reference,
    values_from = total_cnv_per_cell
  ) |>
  # color points this way rather than binning to ensure small bins aren't left out
  dplyr::add_count(`all-immune`, `subset-immune`)

@@

```{r, fig.width = 10}
ggplot(has_cnv_df) +
  # keep in default order so "reference" and "unknown" are grouped
  aes(x = subcluster) +
  geom_bar() +
  facet_wrap(
    vars(normal_reference),
    scales = "free_x"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

```{r fig.width=12}
ggplot(query_has_cnv_df) +
  aes(x = subcluster, fill = forcats::fct_lump(consensus_annotation, 7)) +
  geom_bar() +
  scale_fill_brewer(palette = "Dark2") +
  facet_wrap(
    vars(normal_reference),
    scales = "free_x"
  ) +
  labs(
    title = "Consensus cell types",
    fill = "annotation"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r fig.width=12}
ggplot(query_has_cnv_df) +
  aes(x = subcluster, fill = forcats::fct_lump(ewing_annotation, 7)) +
  geom_bar() +
  scale_fill_brewer(palette = "Set1") +
  facet_wrap(
    vars(normal_reference),
    scales = "free_x"
  ) +
  labs(
    title = "Ewing cell types",
    fill = "annotation"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

```{r fig.width=10}
all_reference_clusters <- has_cnv_df |>
  dplyr::filter(
    normal_reference == "all-immune",
    immune_reference_cell
  )
all_plot <- ggplot(all_reference_clusters) +
  aes(x = subcluster, fill = library_id) +
  geom_bar() +
  labs("All immune reference") +
  theme(legend.position = "bottom")

subset_reference_clusters <- has_cnv_df |>
  # bring in library_id
  dplyr::inner_join(
    infercnv_df |>
      dplyr::select(barcodes, library_id) |>
      unique()
  ) |>
  dplyr::filter(
    normal_reference == "subset-immune",
    subset_reference_cell
  )
subset_plot <- ggplot(subset_reference_clusters) +
  aes(x = subcluster, fill = library_id) +
  geom_bar() +
  labs("Subset immune reference") +
  theme(legend.position = "bottom")

@@

```{r}
# How many CNV per cell per cluster?
ggplot(has_cnv_df) +
  aes(x = subcluster, y = total_cnv_per_cell) +
  geom_point(size = 2) +
  facet_wrap(
    vars(normal_reference),
    scales = "free_x"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
